---
title: "growth_withImageJ"
output: html_document
date: "2025-07-03"
---
### Load libraries
```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(lmtest)
library(car)
install.packages("lmerTest")
library(lmerTest)
install.packages("modelr")
library(modelr)
install.packages("glmmTMB")
library(glmmTMB)#follows lme4 syntax
install.packages("DHARMa")
library(DHARMa)
```


### Load data

```{r}
B_growth_sum <- read_xlsx("/Users/graceleuchtenberger/Library/Mobile Documents/com~apple~CloudDocs/Documents/project-gigas-conditioning-GL/data/sizes/B_2023POGS_small_seed_daily/B_sizes_12172024.xlsx")
B_growth_sum$date <- as.Date(B_growth_sum$date)
B_growth_sum <- B_growth_sum[which(B_growth_sum$date < as.Date("2024-12-17")),]
B_growth_sum$length.mm <- as.numeric(B_growth_sum$length.mm)
B_growth_sum$width.mm <- as.numeric(B_growth_sum$width.mm)
B_growth_sum <- subset(B_growth_sum, select=-c(bag_tag_num,notes, tag_num, duplicate,height.mm ))

D_growth_sum <- read_xlsx("/Users/graceleuchtenberger/Library/Mobile Documents/com~apple~CloudDocs/Documents/project-gigas-conditioning-GL/data/sizes/D_2023POGS_small_seed_weekly/D_sizes_12172024.xlsx", guess_max = 2000)
D_growth_sum$date <- as.Date(D_growth_sum$date)
D_growth_sum <- D_growth_sum[which(D_growth_sum$date < as.Date("2024-12-17")),]
D_growth_sum <- subset(D_growth_sum, select=-c(bag_tag_num,notes, tag_num, duplicate,height.mm ))

Dec_sizes <- read_xlsx("/Users/graceleuchtenberger/Library/Mobile Documents/com~apple~CloudDocs/Documents/project-gigas-conditioning-GL/data/outplanting/Westcott/20241217_images/sizes_12172024.xlsx")
May_sizes <- read_xlsx("/Users/graceleuchtenberger/Library/Mobile Documents/com~apple~CloudDocs/Documents/project-gigas-conditioning-GL/data/outplanting/Westcott/20250513_images/sizes_05132025.xlsx")


```

### Make ImageJ data compatible with early data
```{r}
Dec_sizes <- Dec_sizes %>%
  pivot_wider(names_from = metric,
              values_from = c(measure.mm))

Dec_sizes$width[Dec_sizes$width < 2] <- NA
names(Dec_sizes)[names(Dec_sizes) == "length"] <- "length.mm"
names(Dec_sizes)[names(Dec_sizes) == "width"] <- "width.mm"
Dec_sizes <- subset(Dec_sizes, select=-c(oyster))

May_sizes <- May_sizes %>%
  pivot_wider(names_from = metric,
              values_from = c(measure.mm))

May_sizes$width[May_sizes$width < 2] <- NA

names(May_sizes)[names(May_sizes) == "length"] <- "length.mm"
names(May_sizes)[names(May_sizes) == "width"] <- "width.mm"
May_sizes <- subset(May_sizes, select=-c(oyster))

bag_tag_key <- read_xlsx("/Users/graceleuchtenberger/Library/Mobile Documents/com~apple~CloudDocs/Documents/project-gigas-conditioning-GL/data/outplanting/Westcott/bag_tag_key.xlsx")

B_growth_sum <- left_join(B_growth_sum, bag_tag_key, by = "bag_num")
D_growth_sum  <- left_join(D_growth_sum, bag_tag_key, by = "bag_num")

B_growth_sum <- subset(B_growth_sum, select=-c(bag_num))
D_growth_sum <- subset(D_growth_sum, select=-c(bag_num))


```

### Combine ImageJ and early data
```{r}

growth <- bind_rows(B_growth_sum, D_growth_sum, Dec_sizes, May_sizes)
growth$date <- as.Date(growth$date)
growth$conditioning <- gsub("weekly", "treated", growth$conditioning)
growth$conditioning <- gsub("daily", "treated", growth$conditioning)

growth$predicted <- predict(model, newdata = growth)

```


### Mean sizes
```{r}
means_sizes <- growth %>%
  group_by(effort, conditioning, date) %>%
  summarise(mean_length = mean(length.mm, na.rm = TRUE),
            mean_width = mean(width.mm, na.rm = TRUE),
            stderr_length = sd(length.mm, na.rm = TRUE)/sqrt(length(length.mm)),
            stderr_width = sd(width.mm, na.rm = TRUE)/sqrt(length(width.mm)),
            .groups = 'drop')

```

```{r}
DL <- ggplot(means_sizes, aes(x = date, y = mean_length, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Mean length (mm)") +
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_length-stderr_length, ymax = mean_length+stderr_length), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.x = element_blank(),       
    axis.text.x = element_blank())

DW <- ggplot(means_sizes, aes(x = date, y = mean_width, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Mean width (mm)") +
  theme_classic()+
   scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  labs(color = "Conditioning")+
  geom_errorbar(aes(ymin=mean_width-stderr_width, ymax = mean_width+stderr_width), width=.2,
                 position=position_dodge(0.05))

grid.arrange(DL,DW)
```

```{r}
ggplot(means_sizes, aes(x = date, y = mean_length, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Mean length (mm)") +
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_length-stderr_length, ymax = mean_length+stderr_length), width=.2,
                 position=position_dodge(0.05))
```

## Run after you've predicted volume with the growth model

### Looking at percent change in volume 
```{r}
means_vols <- growth %>%
  group_by(effort, conditioning, date, bag_tag_num) %>%
  summarise(mean_vol = mean(predicted, na.rm = TRUE),
            std_error = sd(predicted, na.rm = TRUE)/sqrt(length(predicted)),
            .groups = 'drop')


means_vols <- means_vols %>%
  group_by(bag_tag_num) %>%
  mutate(initial_vol = first(mean_vol)) %>%
  ungroup()


means_vols <- means_vols %>%
  mutate(perc_change_vol = 100*(mean_vol-initial_vol)/initial_vol ) 
```

```{r}
ggplot(means_vols, aes(x = date, y = mean_vol, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Mean volume ") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_vol-std_error, ymax = mean_vol+std_error), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.x = element_text(size = 20, color = "black"),       
    axis.text.x = element_text(size=18, color = "black"),
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))
```



```{r}
means_perc_vols <- means_vols %>%
  group_by(effort, conditioning, date) %>%
  summarise(mean_perc_vol = mean(perc_change_vol),
            std_error = sd(perc_change_vol)/sqrt(length(perc_change_vol)),
            .groups = 'drop')

means_perc_vols$date <- as.POSIXct(means_perc_vols$date)

ggplot(means_perc_vols, aes(x = date, y = mean_perc_vol, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Percent change in volume") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_perc_vol-std_error, ymax = mean_perc_vol+std_error), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.x = element_text(size = 20, color = "black"),       
    axis.text.x = element_text(size=18, color = "black"),
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))


```

```{r}
library(scales)  # for date_format()

ggplot(means_perc_vols, aes(x = date, y = mean_perc_vol, color = effort)) +
  geom_point() +
  geom_line(aes(linetype = conditioning)) +
  labs(x = "Date",
       y = "Percent change in volume") +
  labs(color = "Conditioning Type", linetype = "Treatment") +
  theme_classic() +
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple", "forestgreen")) +
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed")) +
  geom_errorbar(aes(ymin = mean_perc_vol - std_error, ymax = mean_perc_vol + std_error), 
                width = 0.2,
                position = position_dodge(0.05)) +
  scale_x_datetime(date_labels = "%b %Y", date_breaks = "1 month") +  # <-- add this line
  theme(
    axis.title.x = element_text(size = 20, color = "black"),       
    axis.text.x = element_text(size = 18, color = "black", angle = 45, hjust = 1),  # angled for readability
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size = 18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))

```



```{r}
means_vols %>%
  filter(effort == "D") %>%
  ggplot(aes(x=date, y= perc_change_vol, color = conditioning))+
  geom_point()+
  theme_classic()+
  labs(x = "Date", y = "Percent change volume")+
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 20))
```
```{r}
means_vols %>%
  filter(effort == "B") %>%
  ggplot(aes(x=date, y= perc_change_vol, color = conditioning))+
  geom_point()+
  theme_classic()+
  labs(x = "Date", y = "Percent change volume")+
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 20))
```
```{r}
means_vols %>%
  filter(effort == "B") %>%
  ggplot(aes(x=date, y= mean_vol, color = conditioning))+
  geom_point()+
  theme_classic()+
  labs(x = "Date", y = "Volume")+
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 20))
```
```{r}
means_vols %>%
  filter(effort == "D") %>%
  ggplot(aes(x=date, y= mean_vol, color = conditioning))+
  geom_point()+
  theme_classic()+
  labs(x = "Date", y = "Volume")+
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 20))
```

```{r}
ggplot(means_perc_vols, aes(x = date, y = mean_perc_vol, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Percent change in volume") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  theme(
    axis.title.x = element_text(size = 20, color = "black"),       
    axis.text.x = element_text(size=18, color = "black"),
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))
```

### Percent change in volume model
```{r}
means_vols$date <- as.POSIXct(means_vols$date)
means_vols$bag_tag_num <- as.factor(means_vols$bag_tag_num)
means_vols
model <- lmer(perc_change_vol ~  effort/conditioning *date + (1|bag_tag_num), data = means_vols)
summary(model)
```

```{r}
plot(means_vols$date, means_vols$perc_change_vol)
```

```{r}
means_vols$days <- as.numeric(means_vols$date - min(means_vols$date))
means_vols$days <- means_vols$days/(60*60*24)
model.m<-glmmTMB(perc_change_vol ~  effort/conditioning *days+ (1|bag_tag_num),data = means_vols, family = gaussian()) #glmmTMB
summary(model.m)
car::Anova(model.m,type = "III")
qqPlot(residuals(model.m))

model.b<- lmer(sqrt(perc_change_vol) ~  effort/conditioning *days + (1|bag_num), data = means_vols)
summary(model.b)
qqnorm(resid(model.b))
qqline(resid(model.b))

testDispersion(model.m)
tst_simres <- simulateResiduals(fittedModel = model.m) #DHARMa
plot(tst_simres)

testDispersion(model.b)
tst_simres <- simulateResiduals(fittedModel = model.b) #DHARMa
plot(tst_simres)


```
```{r}
vols_w <- means_vols %>% filter(effort == "D")
vols_d <- means_vols %>% filter(effort == "B")


model.w<-glmmTMB(perc_change_vol ~  conditioning *days+ (1|bag_tag_num),data = vols_w, family = gaussian()) #glmmTMB
summary(model.w)
car::Anova(model.w,type = "III")
qqPlot(residuals(model.w))

testDispersion(model.w)
tst_simres <- simulateResiduals(fittedModel = model.w) #DHARMa
plot(tst_simres)

model.d<-glmmTMB(perc_change_vol ~  conditioning *days+ (1|bag_tag_num),data = vols_d, family = gaussian()) #glmmTMB
summary(model.d)
car::Anova(model.d,type = "III")
qqPlot(residuals(model.d))

testDispersion(model.d)
tst_simres <- simulateResiduals(fittedModel = model.d) #DHARMa
plot(tst_simres)
```

```{r}
hist(means_vols$perc_change_vol)
hist(log(means_vols$perc_change_vol))
hist(sqrt(means_vols$perc_change_vol))
```

#### Pairwise comparisons for percent change volume model
```{r}
install.packages("emmeans")
library(emmeans)
emm_interaction <- emmeans(model.m, ~ effort * conditioning | days)
emm_interaction

# Pairwise comparisons for interaction
pairs(emm_interaction, by = "days", adjust= "tukey")  # Compare effort * conditioning combinations within days


```
### Percent change in length, width, height
```{r}
means_sizes <- growth %>%
  group_by(effort, conditioning, date, bag_tag_num) %>%
  summarise(mean_length = mean(length.mm, na.rm = TRUE),
            mean_width = mean(width.mm, na.rm = TRUE),
            stderr_length = sd(length.mm, na.rm = TRUE)/sqrt(length(length.mm)),
            stderr_width = sd(width.mm, na.rm = TRUE)/sqrt(length(width.mm)),
            .groups = 'drop')

means_sizes <- means_sizes %>%
  group_by(bag_tag_num) %>%
  mutate(initial_length = first(mean_length)) %>%
  mutate(initial_width = first(mean_width)) %>%
  ungroup()


means_sizes <- means_sizes %>%
  mutate(perc_change_length = 100*(mean_length-initial_length)/initial_length ) %>%
  mutate(perc_change_width = 100*(mean_width-initial_width)/initial_width )
```

```{r}
means_sizes %>%
  filter(effort == "D") %>%
  ggplot(aes(x=date, y= perc_change_length, color = conditioning))+
  geom_point()+
  theme_classic()+
  labs(x = "Date", y = "Percent change length")+
  theme(axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.text = element_text(size = 20))
```

```{r}
means_perc_dimension <- means_sizes %>%
  group_by(effort, conditioning, date) %>%
  summarise(mean_perc_length = mean(perc_change_length),
            std_error_length = sd(perc_change_length)/sqrt(length(perc_change_length)),
            mean_perc_width = mean(perc_change_width),
            std_error_width = sd(perc_change_width)/sqrt(length(perc_change_width)),
            .groups = 'drop')

means_perc_dimension$date <- as.POSIXct(means_perc_dimension$date)
```

```{r}
length <- ggplot(means_perc_dimension, aes(x = date, y = mean_perc_length, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Percent change length") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_perc_length-std_error_length, ymax = mean_perc_length+std_error_length), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.x = element_blank(),       
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))

width <- ggplot(means_perc_dimension, aes(x = date, y = mean_perc_width, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Percent change width") +
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  labs(color = "Conditioning Type", linetype = "Treatment")+
  geom_errorbar(aes(ymin=mean_perc_width-std_error_width, ymax = mean_perc_width+std_error_width), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.x = element_blank(),       
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))

grid.arrange(length,width)
```
### Fitting volume models to each bag
```{r}
# first part
#Graph all volume data by bag
#somehow fit models by bag number 
#save coefficients
#graph fits of model to actual values


growth_list <- growth %>%
  group_by(bag_num) %>%
  group_split()

gmodel_list <- lapply(growth_list, function(x) nls(log(volume) ~ a*length.mm^b, data = growth, start = list(a = 6.8, b = 1.1)))



pred1 <- lapply(1:length(growth_list), 
                function(x) predict(growth_list[[x]], 
                                    newdata = growth[[length.mm]], 
                                    type = "response"))


#second part
# use models to predict volumes based on imageJ length data

```

```{r}

growth_split <- split(growth, growth$bag_num)

lapply(growth_split, function(growth) {
  ggplot(growth, aes(x = length.mm, y = log(volume), color = date)) +
    geom_point()+
    geom_line(aes(y = predicted), color = "red", size = 1) + 
    theme_classic()+
    ylim(0,15)+
    xlim(0,90)
})


```


### Model just for length
```{r}
means_sizes$date <- as.Date(means_sizes$date)
means_sizes$days <- as.numeric(means_sizes$date - min(means_sizes$date))
model.m<-glmmTMB(perc_change_length ~  effort/conditioning *days+ (1|bag_tag_num),data = means_sizes, family = gaussian()) #glmmTMB
summary(model.m)
car::Anova(model.m, type = "III")
```


```{r}
means_sizes_big <- growth %>%
  group_by(effort, conditioning, date) %>%
  summarise(mean_length = mean(length.mm, na.rm = TRUE),
            mean_width = mean(width.mm, na.rm = TRUE),
            stderr_length = sd(length.mm, na.rm = TRUE)/sqrt(length(length.mm)),
            stderr_width = sd(width.mm, na.rm = TRUE)/sqrt(length(width.mm)),
            .groups = 'drop')
```

```{r}
ggplot(means_sizes_big, aes(x = date, y = mean_length, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Length") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_length-stderr_length, ymax = mean_length+stderr_length), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))
```

```{r}
ggplot(means_sizes_big, aes(x = date, y = mean_width, color = effort)) +
   geom_point() +
  geom_line(aes(linetype=conditioning))+
  labs(x = "Date",
       y = "Length") +
  labs(color = "Conditioning Type", linetype = "Treatment")+
  theme_classic()+
  scale_color_manual(labels = c("Daily", "Weekly"), values = c("purple","forestgreen"))+
  scale_linetype_manual(labels = c("Control", "Primed"), values = c("solid", "dashed"))+
  geom_errorbar(aes(ymin=mean_width-stderr_width, ymax = mean_width+stderr_width), width=.2,
                 position=position_dodge(0.05))+
  theme(
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    legend.text = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 20, color = "black"))
```



