---
title: Goose Point oyster size analysis
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Set up 

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries. 
```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 

library("ggplot2")
library("tidyverse")
library("lme4")
library("lmerTest")
library("emmeans")
library("car")
```

# Load data 

Read in data files. 
```{r}
data<-read_csv(file="data/outplanting/GoosePoint/growth_GoosePoint.csv")
```

Set data attributes. 
```{r}
data$field_cattle_tag<-factor(data$field_cattle_tag)
```

Set value to 0 for dead or stuck if NA. 
```{r}
data$dead[is.na(data$dead)] <- 0
data$stuck[is.na(data$stuck)] <- 0
```

Add in treatment information 
```{r}
metadata<-read_csv(file="data/outplanting/GoosePoint/bag_list_GoosePoint.csv")
metadata$field_cattle_tag<-factor(metadata$field_cattle_tag)

data$treatment<-metadata$treatment[match(data$field_cattle_tag, metadata$field_cattle_tag)]
```

# Plot data 

Plot a histogram of each data column to identify any outliers. 
```{r}
hist(data$length.mm)
hist(data$width.mm)
hist(data$depth.mm)
```

No obvious outliers. 

Plot a box plot of length, width, and height as a function of bag number, colored by treatment. 

```{r}
plot1<-data%>%
  ggplot(aes(x=field_cattle_tag, y=length.mm, fill = treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot1
```

Plot width. 
```{r}
plot2<-data%>%
  ggplot(aes(x=field_cattle_tag, y=width.mm, fill = treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot2
```

Plot depth 
```{r}
plot3<-data%>%
  ggplot(aes(x=field_cattle_tag, y=depth.mm, fill = treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot3
```

# Generate a metric of volume 

Calculate oyster volume as an ellipsoid. 

```{r}
data$volume <- (4/3) * pi * (data$length.mm/2) * (data$width.mm/2) * (data$depth.mm/2)
```

Look at a histogram. 

```{r}
hist(data$volume)
```

Now plot volume. 

```{r}
plot4<-data%>%
  ggplot(aes(x=field_cattle_tag, y=volume, fill = treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot4
```

Plot summarized by treatment 

```{r}
plot5<-data%>%
  ggplot(aes(x=treatment, y=volume, fill = treatment)) +
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Treatment") + 
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot5
```

Plot summarized by mean and standard error for each bag. 

```{r}
plot6<-data%>%
  group_by(field_cattle_tag, treatment)%>%
  summarise(mean=mean(volume, na.rm=TRUE), se=sd(volume, na.rm=TRUE)/sqrt(length(volume)))%>%
  
  ggplot(aes(x=field_cattle_tag, y=mean, fill = treatment)) +
  geom_point(pch = 21) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0)+
  xlab("Bag") + 
  ylab("Volume")+
  ylim(2000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot6
```

```{r}
plot7<-data%>%
  group_by(treatment)%>%
  summarise(mean=mean(volume, na.rm=TRUE), se=sd(volume, na.rm=TRUE)/sqrt(length(volume)))%>%
  
  ggplot(aes(x=treatment, y=mean, fill = treatment)) +
  geom_point(pch = 21, size=4) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0)+
  xlab("Treatment") + 
  ylab("Volume")+
  ylim(5000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot7
```

Try a violin plot for treatment. 
```{r}
plot8<-data%>%
  ggplot(aes(x=treatment, y=volume, fill = treatment)) +
  geom_violin()+
  xlab("Treatment") + 
  ylab("Volume")+
  #ylim(5000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot8
```

Try a violin plot for bag 
```{r}
plot9<-data%>%
  ggplot(aes(x=field_cattle_tag, y=volume, fill = treatment)) +
  geom_violin()+
  xlab("Bag") + 
  ylab("Volume")+
  #ylim(5000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot9
```

# Run linear mixed models models 

```{r}
model<-lmer(sqrt(volume) ~ treatment + (1|field_cattle_tag:treatment), data=data)
summary(model)
anova(model)

qqPlot(residuals(model))
hist(residuals(model))
```

No significant effect of treatment on volume. Square root transformed to meet assumptions of normality.  

```{r}
model2<-aov(sqrt(volume) ~ treatment*field_cattle_tag, data=data)
summary(model2)

qqPlot(residuals(model2))
```

Field cattle tag is significant, so we need to use the linear mixed model to account for tag.  


