---
title: "Goose Point oyster size analysis"
author: "AS Huffmyer"
date: '2024'
output:
  github_document: default
  github_md: default
editor_options:
  chunk_output_type: console
---

# Set up 

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries. 
```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse') 
if ("ggplot2" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggplot2') 
if ("ggeffects" %in% rownames(installed.packages()) == 'FALSE') install.packages('ggeffects') 

library("ggplot2")
library("tidyverse")
library("lme4")
library("lmerTest")
library("emmeans")
library("car")
library("mgcv")
library("ggeffects")
library("cowplot")
```

# Load data 

Read in data files. 
```{r}
data<-read_csv(file="data/outplanting/GoosePoint/growth_GoosePoint.csv")%>%
  mutate(date=as.factor(date))
```

Set data attributes. 
```{r}
data$field_cattle_tag<-factor(data$field_cattle_tag)
```

Set value to 0 for dead or stuck if NA. 
```{r}
data$dead[is.na(data$dead)] <- 0
data$stuck[is.na(data$stuck)] <- 0
```

Add in treatment information 
```{r}
metadata<-read_csv(file="data/outplanting/GoosePoint/bag_list_GoosePoint.csv")
metadata$field_cattle_tag<-factor(metadata$field_cattle_tag)

data$treatment<-metadata$treatment[match(data$field_cattle_tag, metadata$field_cattle_tag)]
```

# Plot data 

Plot a histogram of each data column to identify any outliers. 
```{r}
hist(data$length.mm)
hist(data$width.mm)
hist(data$depth.mm)
```

No obvious outliers. 

Plot a box plot of length, width, and depth as a function of bag number, colored by treatment. 

```{r}
plot1<-data%>%
  ggplot(aes(x=field_cattle_tag, y=length.mm, fill = treatment)) +
  facet_wrap(~date)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot1
```

Plot width. 
```{r}
plot2<-data%>%
  ggplot(aes(x=field_cattle_tag, y=width.mm, fill = treatment)) +
  facet_wrap(~date)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot2
```

Plot depth 
```{r}
plot3<-data%>%
  ggplot(aes(x=field_cattle_tag, y=depth.mm, fill = treatment)) +
  facet_wrap(~date)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot3
```

# Generate a metric of volume

Generate an equation to predict depth from length and width using 20240909 data and then use this equation to calculate predicted volume from length and width for other dates. 

Calculate oyster volume as an ellipsoid using known. 

```{r}
data$volume <- (4/3) * pi * (data$length.mm/2) * (data$width.mm/2) * (data$depth.mm/2)
```

View the relationship between length, width, and volume. 
```{r}
plot(data$volume ~ data$length.mm)
plot(data$volume ~ data$width.mm)
```
Relationships may be non linear. 

Obtain training data with known length width and depth from 20240909 dataset. 
```{r}
training_data<-data%>%
  filter(date=="20240909")%>%
  filter(!is.na(volume))

hist(training_data$volume)
```

Fit a polynomial regression and a GAM model and compare the fits. 
```{r}
# Fit a polynomial regression model (2nd-degree polynomial)
poly_model <- lm(volume ~ poly(length.mm, 2) + poly(width.mm, 2) + length.mm:width.mm, data = training_data)

# Fit a Generalized Additive Model (GAM) for comparison
gam_model <- gam(volume ~ s(length.mm, bs = "cs") + s(width.mm, bs = "cs"), data = training_data)

# Compare model summaries
summary(poly_model)
summary(gam_model)
```
Poly model adjusted R sq = 0.757
GAM model adjusted R sq = 0.755 

Both models have similar fits. Plot the models. 
```{r}
plot(poly_model)
plot(gam_model)
```

Predict volume using both models and evaluate fits.   
```{r}
# Predict Volume using the trained model
data$Predicted_Volume_GAM <- predict(gam_model, data)
data$Predicted_Volume_Poly <- predict(poly_model, data)
```

Plot predicted vs actual volume for each model. 

```{r}
plot(data$Predicted_Volume_GAM ~ data$volume)
plot(data$Predicted_Volume_Poly ~ data$volume)
```

Relationships look similar for both models. 

Plot GAM relationships
```{r}
# Plot the relationship for visualization
ggplot(training_data, aes(x = length.mm, y = volume)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), color = "blue") +
  ggtitle("GAM Fit for Length vs Volume")

ggplot(training_data, aes(x = width.mm, y = volume)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"), color = "red") +
  ggtitle("GAM Fit for Width vs Volume")
```

Plot Poly relationships
```{r}
# Plot polynomial fit for Length vs. Volume
ggplot(training_data, aes(x = length.mm, y = volume)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Length vs Volume")

# Plot polynomial fit for Width vs. Volume
ggplot(training_data, aes(x = width.mm, y = volume)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Width vs Volume")
```

I am going to select the polynomial model, because the GAM is more affected by higher observations in width. 

View the relationship between known volume and predicted volume using Polynomial regression. 
```{r}
# Plot polynomial fit for Length vs. Volume
ggplot(data%>%filter(date=="20240909"), aes(x = volume, y = Predicted_Volume_Poly)) +
  geom_point() +
  stat_smooth(method = "loess", color = "blue", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Known vs Predicted Volume")
```

Model fit looks good for our approximation. Proceed with using predicted volume calculated by polynomial regression. 

Look for outliers. 
```{r}
hist(data$Predicted_Volume_Poly)
min(data$Predicted_Volume_Poly)
```

Remove outliers and reconstruct polynomial model. 

There are several outliers. Conduct outlier removal and analyze again.  

```{r}
# Fit a polynomial regression model (2nd-degree polynomial)
poly_model <- lm(volume ~ poly(length.mm, 2) + poly(width.mm, 2) + length.mm:width.mm, data = training_data)

# Extract raw residuals
training_data$raw_resid <- residuals(poly_model)

# Standardize residuals
training_data$std_resid <- training_data$raw_resid / sd(training_data$raw_resid)

# Flag potential outliers
outlier_threshold <- 3
training_data$outlier_flag <- abs(training_data$std_resid) > outlier_threshold

# Filter rows flagged as outliers
outliers <- training_data %>% filter(outlier_flag == TRUE)
print(outliers)

# Plot standardized residuals
ggplot(training_data, aes(x = seq_along(std_resid), y = std_resid)) +
  geom_point(aes(color = outlier_flag), size = 2) +
  geom_hline(yintercept = c(-outlier_threshold, outlier_threshold), linetype = "dashed", color = "red") +
  labs(title = "Standardized Residuals with Outliers", x = "Index", y = "Standardized Residual") +
  theme_minimal()
```

Re run the model after removing outliers. 

```{r}
training_data<-training_data%>%
  filter(!outlier_flag==TRUE)
```

```{r}
hist(training_data$volume)
```

Fit a polynomial regression. 
```{r}
# Fit a polynomial regression model (2nd-degree polynomial)
poly_model <- lm(volume ~ poly(length.mm, 2) + poly(width.mm, 2) + length.mm:width.mm, data = training_data)

# Compare model summaries
summary(poly_model)
```
R sq improved to 0.785

```{r}
plot(poly_model)
```

Predict volume using poly model and evaluate fits.   
```{r}
# Predict Volume using the trained model
data$Predicted_Volume_Poly <- predict(poly_model, data)
```

Plot predicted vs actual volume for each model. 

```{r}
plot(data$Predicted_Volume_Poly ~ data$volume)
```

Plot Poly model relationships
```{r}
# Plot polynomial fit for Length vs. Volume
ggplot(training_data, aes(x = length.mm, y = volume)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Length vs Volume")

# Plot polynomial fit for Width vs. Volume
ggplot(training_data, aes(x = width.mm, y = volume)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Width vs Volume")
```

View the relationship between known volume and predicted volume using Polynomial regression. 
```{r}
# Plot polynomial fit for Length vs. Volume
ggplot(data%>%filter(date=="20240909"), aes(x = volume, y = Predicted_Volume_Poly)) +
  geom_point() +
  stat_smooth(method = "loess", color = "blue", se = TRUE) +
  ggtitle("Polynomial Regression (2nd Degree) Fit for Known vs Predicted Volume")
```

Model fit looks good for our approximation. Proceed with using predicted volume calculated by polynomial regression. 

Look for outliers. 
```{r}
hist(data$Predicted_Volume_Poly)
min(data$Predicted_Volume_Poly)
```

There are fewer negative values than the first iteration of the model. 

Remove any observation that resulted in a negative value. 

```{r}
data<-data%>%filter(Predicted_Volume_Poly>0)
hist(data$Predicted_Volume_Poly)
```

Proceed with the calculated volume data.  

# Analyze growth (volume) over time 

## Plot data 

```{r}
plot4<-data%>%
  ggplot(aes(x=field_cattle_tag, y=Predicted_Volume_Poly, fill = treatment)) +
  facet_wrap(~date)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Bag") + 
  theme_classic(); plot4
```

Plot summarized by treatment 

```{r}
plot5<-data%>%
  ggplot(aes(x=treatment, y=Predicted_Volume_Poly, fill = treatment)) +
  facet_wrap(~date)+
  geom_boxplot(width=.5, outlier.shape= NA, position = position_dodge(width = 0.4)) +
  geom_point(pch = 21, position=position_jitterdodge(dodge.width=0.4)) +
  xlab("Treatment") + 
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot5
```

Plot summarized by mean and standard error for each bag. 

```{r}
plot6<-data%>%
  group_by(field_cattle_tag, treatment, date)%>%
  summarise(mean=mean(Predicted_Volume_Poly, na.rm=TRUE), se=sd(Predicted_Volume_Poly, na.rm=TRUE)/sqrt(length(Predicted_Volume_Poly)))%>%
  
  ggplot(aes(x=field_cattle_tag, y=mean, fill = treatment)) +
  facet_wrap(~date)+
  geom_point(pch = 21) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0)+
  xlab("Bag") + 
  ylab("Volume")+
  #ylim(2000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot6
```

```{r}
plot7<-data%>%
  group_by(treatment, date)%>%
  summarise(mean=mean(Predicted_Volume_Poly, na.rm=TRUE), se=sd(Predicted_Volume_Poly, na.rm=TRUE)/sqrt(length(Predicted_Volume_Poly)))%>%
  
  ggplot(aes(x=treatment, y=mean, fill = treatment)) +
  facet_wrap(~date)+
  geom_point(pch = 21, size=4) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0)+
  xlab("Treatment") + 
  ylab("Volume")+
  #ylim(5000, 15000)+
  theme_classic() + 
  theme(axis.text.x=element_text(angle=45, vjust=1, hjust=1)); plot7
```

Sizes vary between treatments at the start of the outplant. We will examine growth using models to calculate slopes of change in volume over time.  

## Run linear mixed models models 

```{r}
model<-lmer(sqrt(Predicted_Volume_Poly) ~ treatment * date + (1|field_cattle_tag:treatment), data=data)

summary(model)
anova(model)

qqPlot(residuals(model))
hist(residuals(model))
```

Different sizes by treatment and treatment by date. 

There are several outliers. Conduct outlier removal and analyze again.  

```{r}
# Extract raw residuals
data$raw_resid <- residuals(model)

# Standardize residuals
data$std_resid <- data$raw_resid / sd(data$raw_resid)

# Flag potential outliers
outlier_threshold <- 3
data$outlier_flag <- abs(data$std_resid) > outlier_threshold

# Filter rows flagged as outliers
outliers <- data %>% filter(outlier_flag == TRUE)
print(outliers)

# Plot standardized residuals
ggplot(data, aes(x = seq_along(std_resid), y = std_resid)) +
  geom_point(aes(color = outlier_flag), size = 2) +
  geom_hline(yintercept = c(-outlier_threshold, outlier_threshold), linetype = "dashed", color = "red") +
  labs(title = "Standardized Residuals with Outliers", x = "Index", y = "Standardized Residual") +
  theme_minimal()
```

Remove identified outliers. 
```{r}
data<-data%>%
  filter(!outlier_flag==TRUE)
```

Analyze again. 

```{r}
model<-lmer(sqrt(Predicted_Volume_Poly) ~ treatment * date + (1|field_cattle_tag:treatment), data=data)

summary(model)
anova(model)

qqPlot(residuals(model))
hist(residuals(model))
```

Residuals are much better.  

# Analyze growth rates over time 

```{r}
# Fit a mixed-effects model to estimate growth rate
growth_model <- lmer(Predicted_Volume_Poly ~ date * treatment + (1|treatment:field_cattle_tag), data = data)

# Summarize the model
summary(growth_model)
anova(growth_model)

# Extract fixed effects (overall growth rate) and random effects (bag-specific deviations)
fixed_effects <- fixef(growth_model)
random_effects <- ranef(growth_model)$`treatment:field_cattle_tag`

# Plot growth trajectories
ggplot(data, aes(x = date, y = Predicted_Volume_Poly, color = factor(treatment), group=treatment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Oyster Growth Over Time",
       x = "Date",
       y = "Volume",
       color = "Treatment")

#plot model trajectories 
ggpred <- ggpredict(growth_model, terms = c("date", "treatment"))

ggplot(ggpred, aes(x = x, y = predicted, color = group, fill=group)) +
  geom_line(aes(group=group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, group=group), alpha = 0.2) +
  theme_classic() +
  labs(title = "Modeled Oyster Growth Over Time",
       x = "Date",
       y = "Volume",
       color = "Treatment", 
       fill = "Treatment")
```

Analyze each effort separately. 

Weekly fresh water group:  
```{r}
fresh_data<-data[grepl("Fresh Water", data$treatment), ]

# Fit a mixed-effects model to estimate growth rate
growth_model_FW <- lmer(Predicted_Volume_Poly ~ date * treatment + (1|treatment:field_cattle_tag), data = fresh_data)

# Summarize the model
summary(growth_model_FW)
anova(growth_model_FW)

# Extract fixed effects (overall growth rate) and random effects (bag-specific deviations)
fixed_effects <- fixef(growth_model_FW)
random_effects <- ranef(growth_model_FW)$`treatment:field_cattle_tag`

# Plot growth trajectories
ggplot(fresh_data, aes(x = date, y = Predicted_Volume_Poly, color = factor(treatment))) +
  geom_point() +
  geom_smooth(aes(group=treatment), method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Oyster Growth Over Time",
       x = "Date",
       y = "Volume",
       color = "Treatment")

#plot model trajectories 
ggpred <- ggpredict(growth_model_FW, terms = c("date", "treatment"))

plot1<-ggplot(ggpred, aes(x = x, y = predicted, color = group, fill=group, group=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  theme_classic() +
  
  labs(title = "Modeled Oyster Growth Over Time (Fresh Water)",
       x = "Date",
       y = "Predicted Volume",
       color = "Treatment", 
       fill = "Treatment");plot1
```

Treated are very slightly but significantly larger than controls. 

Weekly temperature group:  
```{r}
temp_data<-data[grepl("Temperature", data$treatment), ]

# Fit a mixed-effects model to estimate growth rate
growth_model_TEMP <- lmer(Predicted_Volume_Poly ~ date * treatment + (1|treatment:field_cattle_tag), data = temp_data)

# Summarize the model
summary(growth_model_TEMP)
anova(growth_model_TEMP)

# Extract fixed effects (overall growth rate) and random effects (bag-specific deviations)
fixed_effects <- fixef(growth_model_TEMP)
random_effects <- ranef(growth_model_TEMP)$`treatment:field_cattle_tag`

# Plot growth trajectories
ggplot(temp_data, aes(x = date, y = Predicted_Volume_Poly, color = factor(treatment), group=treatment)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Oyster Growth Over Time",
       x = "Date",
       y = "Volume",
       color = "Treatment")

#plot model trajectories 
ggpred <- ggpredict(growth_model_TEMP, terms = c("date", "treatment"))

plot2<-ggplot(ggpred, aes(x = x, y = predicted, color = group, fill=group, group=group)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  theme_classic() +
  
  labs(title = "Modeled Oyster Growth Over Time (Temperature)",
       x = "Date",
       y = "Predicted Volume",
       color = "Treatment", 
       fill = "Treatment");plot2
```

No clear difference between control and treated in the temperature experiment. 

```{r}
plots<-plot_grid(plot1, plot2, ncol=2, nrow=1);plots

ggsave(filename="figures/growth/goose_point/growth.png", plots, width=14, height=6)
```





