---
title: "Viewing logger files"
author: "Ariana S Huffmyer"
date: "2024"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo loggers. 

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
```

# Point Whitney 

## Read in files 

Read in Hobo Tidbit files  
```{r}
tidbit.files <- list.files(path="data/environmental/loggers/point_whitney/for-analysis", pattern="*.csv", full.names=TRUE)

#will need to change to HST time 
tidbits <- setNames(tidbit.files, tidbit.files) %>%
    map_dfr(read_csv, .id = "logger") %>%
    rename(DateTime=`Date-Time (PDT)`, temp.C=`Temp   (°C)`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

#format date for trays logger, formatted differently 
trays<-tidbits%>%
  filter(logger=="22023395")

other<-tidbits%>%
  filter(!logger=="22023395")

trays$DateTime<-as.POSIXct(trays$DateTime, format="%m/%d/%y %H:%M")
other$DateTime<-as.POSIXct(other$DateTime, format="%m/%d/%Y %H:%M:%S")

tidbits<-rbind(trays,other)
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(tidbits$logger))
```

There are 4 loggers as expected.  

Add in group information. 

```{r}
tidbits <- tidbits %>%
  mutate(group = case_when(
    logger == "22023395" ~ "trays",
    logger == "22023398" ~ "outdoor",
    logger == "22023406" ~ "broodstock",
    logger == "22023410" ~ "LCO",
    TRUE ~ "other"  # Default for other values
  ))

levels(as.factor(tidbits$group))
```

Group information is now added.  

## Plot data 

All groups 
```{r}
temp_plot<-tidbits%>%
  
  ggplot(aes(x=DateTime, y=temp.C, colour=group, fill=group))+
  geom_point(alpha=0.7)+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  ylim(8,32)+
  theme_classic(); temp_plot

ggsave(temp_plot, filename="figures/environmental/point-whitney-loggers.png", width=6, height=4)
```

On August 8th trays were moved to a different stack, which is why the temperature shifted to be lower.  

Overall, all groups in outdoor trays (efforts B/E/D, LCO, and broodstock) experienced very similar temperatures. Trays with 2024 seed were cooler. Summer temps peaked in late July around 26-27°C. 

3-holer 
```{r}
temp_plot<-tidbits%>%
  filter(group %in% c("broodstock", "LCO", "outdoor"))%>%
  filter(temp.C>15)%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour="gray")+
  geom_smooth(colour="black")+
  ylab("Temperature (°C)")+
  xlab("Date")+
  ggtitle("3-holer tanks")+
  ylim(8,32)+
  theme_classic(); temp_plot

ggsave(temp_plot, filename="figures/environmental/3-holer-tanks-PW.png", width=6, height=4)
```

Seed trays 
```{r}
temp_plot<-tidbits%>%
  filter(group %in% c("trays"))%>%
  #filter(temp.C>15)%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour="gray")+
  geom_smooth(colour="black")+
  ylab("Temperature (°C)")+
  xlab("Date")+
  ggtitle("Seed trays")+
  geom_vline(xintercept=as.POSIXct("2024-08-05 15:45:00"), linetype="dashed", colour="black")+
  geom_text(x=as.POSIXct("2024-07-01 15:45:00"), y=22, label="left stack")+
  geom_text(x=as.POSIXct("2024-09-01 15:45:00"), y=22, label="right stack")+
  ylim(8,25)+
  theme_classic(); temp_plot

ggsave(temp_plot, filename="figures/environmental/seed-trays-PW.png", width=6, height=4)
```



# Westcott 

Read in Hobo Tidbit files  
```{r}
westcott.files <- list.files(path="data/environmental/loggers/westcott/westcottloggers_20250513/", pattern="*.xlsx", full.names=TRUE)

#will need to change to HST time 
westcott <- setNames(westcott.files, westcott.files) %>%
    map_dfr(read_xlsx, .id = "logger") %>%
    rename(DateTime=`Date-Time (PDT)`, temp.C=`Temperature , °C`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

westcott$DateTime<-as.POSIXct(westcott$DateTime, format="%m/%d/%Y %H:%M:%S")
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(westcott$logger))
```

There are 5 loggers as expected.  

Add in group information. 

```{r}
westcott <- westcott %>%
  mutate(group = c("Westcott"))

levels(as.factor(westcott$group))
```

Group information is now added.  

## Plot data 

All loggers 
```{r}
westcott_plot<-westcott%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour="darkgray", fill="darkgray")+
  geom_smooth(colour="black", se=FALSE)+
  ylab("Temperature (°C)")+
  xlab("Date Time")+
  ylim(-5,50)+
  theme_classic(); westcott_plot

ggsave(westcott_plot, filename="figures/environmental/westcott-loggers.png", width=10, height=6)
```

Overall, loggers at Westcott show huge swings in temperature with maximum temperatures above 40 and low temperatures at approximately 10C. 

# Goose Point 

Read in Hobo Tidbit files  
```{r}
goose.files <- list.files(path="data/environmental/loggers/goose_point", pattern="*.xlsx", full.names=TRUE)

goose <- setNames(goose.files, goose.files) %>%
    map_dfr(read_excel, .id = "logger") %>%
    rename(DateTime=`Date-Time (PDT)`, temp.C=`Temp , °C`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

#export data 
goose%>%
  write_csv("output/environmental/goose-point-temperature-data.csv")

goose$DateTime<-as.POSIXct(goose$DateTime, format="%m/%d/%y %H:%M")
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(goose$logger))
```

There are 4 loggers as expected.  

Add in group information. 

```{r}
goose <- goose %>%
  mutate(group = c("Goose Point"))

levels(as.factor(goose$group))
```

Group information is now added.  

Remove temperature data from read out days (6/24, 9/9, 12/11)

```{r}
# Dates to filter out
dates_to_filter <- as.Date(c("2024-06-19", "2024-06-20", "2024-06-21", "2024-06-22", "2024-06-23", "2024-06-24", "2024-06-25", "2024-09-09", "2024-09-10", "2024-12-11", "2025-05-20"))

# Filter out rows with specified dates
goose <- goose %>%
  mutate(day=as.Date(DateTime))%>%
  filter(!day %in% dates_to_filter)

```

## Plot data 

All loggers 
```{r}
goose_plot<-goose%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour="darkgray")+
  geom_smooth(colour="black", se=FALSE)+
  ylab("Temperature (°C)")+
  xlab("Date")+
  ylim(-5,40)+
  theme_classic(); goose_plot

ggsave(goose_plot, filename="figures/environmental/goose-loggers.png", width=10, height=6)
```

Plot amount of daily variability over a 24 hour cycle. 

```{r}
#calculate min and max of each day and plot this difference over time 
daily_goose<-goose%>%
  group_by(logger, day, group)%>%
  summarise(min=min(temp.C), max=max(temp.C))%>%
  mutate(diff=max-min)
```

Plot daily range over time. 

```{r}
goose_plot2<-daily_goose%>%
  
  ggplot(aes(x=day, y=diff))+
  geom_point(alpha=0.7, colour="darkgray")+
  geom_smooth(colour="black", se=FALSE)+
  ylab("Daily Temperature Range (°C)")+
  xlab("Date")+
  ylim(-5,20)+
  theme_classic(); goose_plot2

ggsave(goose_plot2, filename="figures/environmental/goose-loggers_daily-range.png", width=8, height=6)
```




# Sequim

Read in Hobo file  
```{r}
sequim.files <- list.files(path="data/environmental/loggers/sequim", pattern="*.csv", full.names=TRUE)

sequim <- setNames(sequim.files, sequim.files) %>%
    map_dfr(read_csv, .id = "logger") %>%
    rename(DateTime=`Date-Time (PDT)`, temp.C=`Temp , °C`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

#export data 
sequim%>%
  write_csv("output/environmental/sequim-temperature-data.csv")

sequim$DateTime<-as.POSIXct(sequim$DateTime, format="%m/%d/%y %H:%M")
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(sequim$logger))
```

Add in group information. 

```{r}
sequim <- sequim %>%
  mutate(group = c("Sequim"))

levels(as.factor(sequim$group))
```

Group information is now added.  

Remove temperature data from read out days. 

```{r}
# Dates to filter out
dates_to_filter <- as.Date(c("2024-04-03"))

# Filter out rows with specified dates
sequim <- sequim %>%
  mutate(day=as.Date(DateTime))%>%
  filter(!day < dates_to_filter)

```

## Plot data 

All loggers 
```{r}
sequim_plot<-sequim%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour="darkgray")+
  geom_smooth(colour="black", se=FALSE)+
  ylab("Temperature (°C)")+
  xlab("Date")+
  #ylim(-5,40)+
  theme_classic(); sequim_plot

ggsave(sequim_plot, filename="figures/environmental/sequim-loggers.png", width=10, height=6)
```

Plot amount of daily variability over a 24 hour cycle. 

```{r}
#calculate min and max of each day and plot this difference over time 
daily_sequim<-sequim%>%
  group_by(logger, day, group)%>%
  summarise(min=min(temp.C), max=max(temp.C))%>%
  mutate(diff=max-min)
```

Plot daily range over time. 

```{r}
sequim_plot2<-daily_sequim%>%
  
  ggplot(aes(x=day, y=diff))+
  geom_point(alpha=0.7, colour="darkgray")+
  geom_smooth(colour="black", se=FALSE)+
  ylab("Daily Temperature Range (°C)")+
  xlab("Date")+
  #ylim(-5,20)+
  theme_classic(); sequim_plot2

ggsave(sequim_plot2, filename="figures/environmental/sequim-loggers_daily-range.png", width=8, height=6)
```



