---
title: 10K seed resazurin survival analysis
author: "AS Huffmyer"
date: '2024'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Set up 

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries. 
```{r}
library(tidyverse)
library(ggplot2)
library(survival)
library(readxl)
library(ggsurvfit)
library(gtsummary)
library(cardx)
library(cowplot)
```

# Read data 

Read in data. 
```{r}
data<-read_excel("data/survival/lab_survival/lab_survival.xlsx")
```

Turn into long format. 
```{r}
data<-data%>%
  pivot_longer(names_to="time", values_to="status", cols=`0`:`24`)%>%
  mutate(time=as.numeric(time))
```

Create data frames for each effort. 
```{r}
effort_d<-data%>%filter(effort=="Effort D")
effort_b<-data%>%filter(effort=="Effort B")
effort_e<-data%>%filter(effort=="Effort E")
```

Create data frames with control and high temperature subsets. 
```{r}
control_d<-effort_d%>%filter(temperature=="18C")
stress_d<-effort_d%>%filter(temperature=="46C")

control_b<-effort_b%>%filter(temperature=="18C")
stress_b<-effort_b%>%filter(temperature=="46C")

control_e<-effort_e%>%filter(temperature=="18C")
stress_e<-effort_e%>%filter(temperature=="46C")
```


# Effort D (Weekly temperature stress) 

## Generate Kaplan Meier survival curves 

### Control temperature 18C

```{r}
s1 <- survfit(Surv(time, status) ~ treatment, data = control_d)
str(s1)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = control_d) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = control_d), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = control_d)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = control_d)

coxph(Surv(time, status) ~ treatment, data = control_d) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate Kaplan Meier survival curves 

### Stress temperature 42C

```{r}
s2 <- survfit(Surv(time, status) ~ treatment, data = stress_d)
str(s2)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = stress_d) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = stress_d), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = stress_d)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = stress_d)

coxph(Surv(time, status) ~ treatment, data = stress_d) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate plots 
 
Set theme. 
```{r}
my_theme<-theme_classic()
```

Control
```{r}
plot1<-survfit2(Surv(time, status) ~ treatment, data = control_d) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title="Effort D (Weekly Temp) 18°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p>0.05")+
  #geom_text(x=10, y=0.2, label="Cox PH p=0.8")+
  theme(legend.position="none")
```

Stress
```{r}
plot2<-survfit2(Surv(time, status) ~ treatment, data = stress_d) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title = "Effort D (Weekly Temp) 46°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p>0.05")+
  #geom_text(x=10, y=0.15, label="Temperature vs control p=0.002")+
  #geom_text(x=10, y=0.10, label="All others vs control p>0.05")+
  theme(legend.position="right")
```

Assemble plot
```{r}
plots<-plot_grid(plot1, plot2, rel_widths=c(0.65,1), ncol=2)

ggsave(plots, filename="figures/survival/KMcurves_EffortD.png", width=10, height=4)
```




# Effort B (Daily temperature stress) 

## Generate Kaplan Meier survival curves 

### Control temperature 18C

```{r}
s1 <- survfit(Surv(time, status) ~ treatment, data = control_b)
str(s1)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = control_b) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = control_b), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = control_b)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = control_b)

coxph(Surv(time, status) ~ treatment, data = control_b) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate Kaplan Meier survival curves 

### Stress temperature 42C

```{r}
s2 <- survfit(Surv(time, status) ~ treatment, data = stress_b)
str(s2)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = stress_b) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = stress_b), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = stress_b)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = stress_b)

coxph(Surv(time, status) ~ treatment, data = stress_b) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate plots 
 
Set theme. 
```{r}
my_theme<-theme_classic()
```

Control
```{r}
plot1<-survfit2(Surv(time, status) ~ treatment, data = control_b) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title="Effort B (Daily Temp) 18°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p>0.05")+
  #geom_text(x=10, y=0.2, label="Cox PH p=0.8")+
  theme(legend.position="none")
```

Stress
```{r}
plot2<-survfit2(Surv(time, status) ~ treatment, data = stress_b) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title = "Effort B (Daily Temp) 46°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p>0.05")+
  #geom_text(x=10, y=0.15, label="Temperature vs control p=0.002")+
  #geom_text(x=10, y=0.10, label="All others vs control p>0.05")+
  theme(legend.position="right")
```

Assemble plot
```{r}
plots<-plot_grid(plot1, plot2, rel_widths=c(0.65,1), ncol=2)

ggsave(plots, filename="figures/survival/KMcurves_EffortB.png", width=10, height=4)
```



# Effort E (Weekly temperature or fresh water stress) 

## Generate Kaplan Meier survival curves 

### Control temperature 18C

```{r}
s1 <- survfit(Surv(time, status) ~ treatment, data = control_e)
str(s1)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = control_e) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = control_e), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = control_e)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = control_e)

coxph(Surv(time, status) ~ treatment, data = control_e) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate Kaplan Meier survival curves 

### Stress temperature 46C

```{r}
s2 <- survfit(Surv(time, status) ~ treatment, data = stress_e)
str(s2)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ treatment+temperature, data = stress_e) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ treatment, data = stress_e), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ treatment, data = stress_e)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ treatment, data = stress_e)

coxph(Surv(time, status) ~ treatment, data = stress_e) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate plots 
 
Set theme. 
```{r}
my_theme<-theme_classic()
```

Control
```{r}
plot1<-survfit2(Surv(time, status) ~ treatment, data = control_e) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title="Effort E (Weekly Temp or Fresh) 18°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p>0.05")+
  #geom_text(x=10, y=0.2, label="Cox PH p=0.8")+
  theme(legend.position="none")
```

Stress
```{r}
plot2<-survfit2(Surv(time, status) ~ treatment, data = stress_e) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title = "Effort E (Weekly Temp or Fresh) 46°C"
  )+
  ylim(0,1)+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p=0.001")+
  #geom_text(x=10, y=0.15, label="Temperature vs control p=0.002")+
  #geom_text(x=10, y=0.10, label="All others vs control p>0.05")+
  theme(legend.position="right")
```

Assemble plot
```{r}
plots<-plot_grid(plot1, plot2, rel_widths=c(0.65,1), ncol=2)

ggsave(plots, filename="figures/survival/KMcurves_EffortE.png", width=10, height=4)
```



