---
title: "Viewing logger files"
author: "Ariana S Huffmyer"
date: "2026"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

This script reads and plots environmental data from Hobo loggers. 

# Set Up    

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(stringr)
library(readxl)
library(purrr)
library(lubridate)
library(ggplot2)
library(seacarb)
library(broom)
library(cowplot)
library(slider)
```

# Westcott Repeat Priming experiment

Read in Hobo Tidbit files  
```{r}
westcott.files <- list.files(path="data/westcott-repeat-priming/loggers", pattern="*.xlsx", full.names=TRUE)

#will need to change to HST time 
westcott <- setNames(westcott.files, westcott.files) %>%
    map_dfr(read_xlsx, .id = "logger") %>%
    rename(DateTime=`Date-Time (PDT)`, temp.C=`Temperature , °C`)%>%
    select(!`#`)%>%
    select(logger, DateTime, temp.C)%>%
    mutate(logger=sub(".*/", "", logger))%>% #remove the path name
    mutate(logger=sub("_.*", "", logger))%>% #keep only the serial number 
    filter(!temp.C=="NA")

westcott$DateTime<-as.POSIXct(westcott$DateTime, format="%m/%d/%Y %H:%M:%S")

#export data 
westcott%>%
  write_csv("output/westcott-repeat-priming/westcott-temperature-data.csv")
```

List the logger serial numbers in this dataframe.  
```{r}
levels(as.factor(westcott$logger))
```

There are 2 loggers as expected. These loggers were with the point whitney oysters during priming and are with all oysters at westcott. 

Add in group information. 

```{r}
westcott <- westcott %>%
  mutate(group = c("Westcott"))

levels(as.factor(westcott$group))
```

Group information is now added.  

## Plot data 

All loggers 
```{r}
westcott_plot<-westcott%>%
  #filter(DateTime > "2026-07-17 00:00:00")%>%
  
  ggplot(aes(x=DateTime, y=temp.C))+
  geom_point(alpha=0.7, colour ="gray")+
  ylab("Temperature (°C)")+
  xlab("Date")+
  scale_x_datetime(
    date_breaks = "1 months",
    date_labels = "%b %Y"
  ) +
  ylim(-5,50)+
  theme_classic()+
  theme(axis.text.y = element_text(size = 14, color= "black"),
        axis.title.y = element_text(size = 14, color= "black"),
        axis.text.x = element_blank(),
        axis.title.x = element_blank());westcott_plot

#ggsave(westcott_plot, filename="figures/environmental/westcott-loggers.png", width=10, height=6)
```

