---
title: Westcott lab survival analysis
author: "AS Huffmyer"
date: '2025'
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
  pdf_document:
    keep_tex: yes
editor_options: 
  chunk_output_type: console
---

# Set up 

Set up workspace, set options, and load required packages.    
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries. 
```{r}
library(tidyverse)
library(ggplot2)
library(survival)
library(readxl)
library(ggsurvfit)
library(gtsummary)
library(cardx)
library(cowplot)
```

# Read in data 

Read in data. 
```{r}
data<-read_csv("data/survival/survival_testing_Dec25/survival_westcott_lab.csv")
```

Turn into long format. 
```{r}
data<-data%>%
  pivot_longer(names_to="time", values_to="status", cols=`0`:`50`)%>%
  mutate(time=as.numeric(time))
```

Create data frames for each experiment 
```{r}
daily<-data%>%filter(experiment=="Daily")
weekly<-data%>%filter(experiment=="Weekly")
```

Create data frames with control and high temperature subsets. 
```{r}
control_daily<-daily%>%filter(test_temperature=="22C")
stress_daily<-daily%>%filter(test_temperature=="33C")

control_weekly<-weekly%>%filter(test_temperature=="22C")
stress_weekly<-weekly%>%filter(test_temperature=="33C")

control<-data%>%filter(test_temperature=="22C")
stress<-data%>%filter(test_temperature=="33C")
```


# Daily temperature hardening experiment  

## Generate Kaplan Meier survival curves - control

### Control temperature 22C

```{r}
s1 <- survival::survfit(Surv(time, status) ~ treatment, data = control_daily)
str(s1)
```

Plot the survival function
```{r}
ggsurvfit::survfit2(Surv(time, status) ~ treatment+test_temperature, data = control_daily) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survival::survfit(Surv(time, status) ~ treatment, data = control_daily), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survival::survdiff(Surv(time, status) ~ treatment, data = control_daily)
```

Analyze again with a Cox proportional hazards model. 
```{r}
survival::coxph(Surv(time, status) ~ treatment, data = control_daily)

survival::coxph(Surv(time, status) ~ treatment, data = control_daily) %>% 
  tbl_regression(exp = TRUE) 
```

100% survival 

## Generate Kaplan Meier survival curves - high

### Stress temperature 33C

```{r}
s2 <- survival::survfit(Surv(time, status) ~ treatment, data = stress_daily)
str(s2)
```

Plot the survival function
```{r}
ggsurvfit::survfit2(Surv(time, status) ~ treatment+test_temperature, data = stress_daily) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survival::survfit(Surv(time, status) ~ treatment, data = stress_daily), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survival::survdiff(Surv(time, status) ~ treatment, data = stress_daily)
```

Analyze again with a Cox proportional hazards model. 
```{r}
survival::coxph(Surv(time, status) ~ treatment, data = stress_daily)

survival::coxph(Surv(time, status) ~ treatment, data = stress_daily) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate plots 
 
Set theme. 
```{r}
my_theme<-theme_classic()+
  theme(
    axis.title.y = element_text(size = 20, color = "black"),
    axis.text.y = element_text(size=18, color = "black"),
    axis.text.x = element_text(size=18, color = "black"),
    axis.title.x =element_text(size = 20, color = "black"),
    legend.position = "none")
```

Control
```{r}
options("ggsurvfit.switch-color-linetype" = TRUE)
plot1<-ggsurvfit::survfit2(Surv(time, status) ~ treatment, data = control_daily) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title="Daily Hardening 22°C"
  )+
  ylim(0,1)+
  my_theme+
  scale_linetype_manual(values=c("solid", "dashed"))+
  geom_text(x=10, y=0.2, label="Cox PH p=1", color="black")+
  #geom_text(x=10, y=0.2, label="Cox PH p=0.8")+
  theme(legend.position="none")
plot1
```

Stress
```{r}
options("ggsurvfit.switch-color-linetype" = TRUE)
plot2<-ggsurvfit::survfit2(Surv(time, status) ~ treatment, data = stress_daily) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title = "Daily Hardening 33°C"
  )+
  ylim(0,1)+
  scale_linetype_manual(values=c("solid", "dashed"))+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p=0.492", color="black")+
  #geom_text(x=10, y=0.15, label="Temperature vs control p=0.002")+
  #geom_text(x=10, y=0.10, label="All others vs control p>0.05")+
  theme(legend.position="right")
plot2
```

Assemble plot
```{r}
plots<-plot_grid(plot1, plot2, rel_widths=c(0.75,1), ncol=2)

ggsave(plots, filename="figures/survival/westcott_lab_survival/daily-hardening_updated.png", width=8, height=4)
```

# Weekly temperature hardening experiment  

## Generate Kaplan Meier survival curves - control 

### Control temperature 22C

```{r}
s1 <- survival::survfit(Surv(time, status) ~ treatment, data = control_weekly)
str(s1)
```

Plot the survival function
```{r}
ggsurvfit::survfit2(Surv(time, status) ~ treatment+test_temperature, data = control_weekly) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survival::survfit(Surv(time, status) ~ treatment, data = control_weekly), times = 24)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survival::survdiff(Surv(time, status) ~ treatment, data = control_weekly)
```

Analyze again with a Cox proportional hazards model. 
```{r}
survival::coxph(Surv(time, status) ~ treatment, data = control_weekly)

survival::coxph(Surv(time, status) ~ treatment, data = control_weekly) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate Kaplan Meier survival curves - high 

### Stress temperature 33C

```{r}
s2 <- survival::survfit(Surv(time, status) ~ treatment, data = stress_weekly)
str(s2)
```

Plot the survival function
```{r}
ggsurvfit::survfit2(Surv(time, status) ~ treatment+test_temperature, data = stress_weekly) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survival::survfit(Surv(time, status) ~ treatment, data = stress_weekly), times = 27)
summary(survival::survfit(Surv(time, status) ~ treatment, data = stress_weekly), times = 31)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survival::survdiff(Surv(time, status) ~ treatment, data = stress_weekly)
```

Analyze again with a Cox proportional hazards model. 
```{r}
survival::coxph(Surv(time, status) ~ treatment, data = stress_weekly)

survival::coxph(Surv(time, status) ~ treatment, data = stress_weekly) %>% 
  tbl_regression(exp = TRUE) 
```

## Generate plots 

Control
```{r}
options("ggsurvfit.switch-color-linetype" = TRUE)
plot1<-ggsurvfit::survfit2(Surv(time, status) ~ treatment, data = control_weekly) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title="Weekly Hardening 22°C"
  )+
  ylim(0,1)+
  scale_linetype_manual(values=c("solid", "dashed"))+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p=1", color="black")+
  #geom_text(x=10, y=0.2, label="Cox PH p=0.8")+
  theme(legend.position="none")
plot1
```

Stress
```{r}
options("ggsurvfit.switch-color-linetype" = TRUE)
plot2<-ggsurvfit::survfit2(Surv(time, status) ~ treatment, data = stress_weekly) %>% 
  ggsurvfit::ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability",
    title = "Weekly Hardening 33°C"
  )+
  ylim(0,1)+
  scale_linetype_manual(values=c("solid", "dashed"))+
  my_theme+
  geom_text(x=10, y=0.2, label="Cox PH p=0.029")+
  #geom_text(x=10, y=0.15, label="Temperature vs control p=0.002")+
  #geom_text(x=10, y=0.10, label="All others vs control p>0.05")+
  theme(legend.position="right")
plot2
```

Assemble plot
```{r}
plots<-plot_grid(plot1, plot2, rel_widths=c(0.75,1), ncol=2)

ggsave(plots, filename="figures/survival/westcott_lab_survival/weekly-hardening-updated.png", width=8, height=4)
```





# Daily vs Weekly comparison

### Stress temperature 33C

```{r}
s7 <- survfit(Surv(time, status) ~ group, data = stress)
str(s7)
```

Plot the survival function
```{r}
survfit2(Surv(time, status) ~ group, data = stress) %>% 
  ggsurvfit() +
  labs(
    x = "Hours",
    y = "Survival probability"
  )
```

Estimate survival at 24 hours. 
```{r}
summary(survfit(Surv(time, status) ~ group, data = stress), times = 27)
summary(survfit(Surv(time, status) ~ group, data = stress), times = 31)
```

Use a log rank model to determine statistical differences in curves. 
```{r}
survdiff(Surv(time, status) ~ group, data = stress)
```

Analyze again with a Cox proportional hazards model. 
```{r}
coxph(Surv(time, status) ~ group, data = stress)

coxph(Surv(time, status) ~ group, data = stress) %>% 
  tbl_regression(exp = TRUE) 
```
